Graphical Tests for Power Comparison of Competing Designs

### 2012-03-30 experiment-two.tex
some more on confidence levels in experiment 2

The two highly summarized designs benefit the most from an increase in sample size: both density charts and boxplots profit significantly more than histograms and dotplots. This might, again, be a hint towards a trade-off between visual complexity and amount of information shown as basis for decisions. 
\begin{table}[ht]
\begin{center}
\resizebox{\linewidth}{!} {
\begin{tabular}{lrrrrrl}
  \hline
& & Estimate & Error & $t$-value & \multicolumn{2}{l}{approx $p$-value} \\   \hline
\\[-8pt]
\bf design& Intercept & -3.11 & 0.41 & -7.61 & 0.00 & ***\\ 
&  density & 0.05 & 0.57 & 0.09 & 0.93 \\ 
 & dotplot & 0.46 & 0.57 & 0.81 & 0.42 \\ 
  &histogram & 0.06 & 0.58 & 0.11 & 0.91 \\ [2pt]
\multicolumn{2}{l}{\bf covariates}\\
&  d & 1.77 & 0.34 & 5.18 & 0.00 & ***\\ [1pt]
&   n1$_{100}$ & 1.84 & 0.19 & 9.73 & 0.00 & *** \\ [1pt]
&  ratio & 0.84 & 0.33 & 2.56 & 0.01 & ** \\ [2pt]
\multicolumn{2}{l}{\bf interactions}\\
&    density:d & -0.60 & 0.47 & -1.27 & 0.20 \\ 
&    dotplot:d & 0.03 & 0.46 & 0.07 & 0.95 \\ 
&    histogram:d & -0.28 & 0.48 & -0.59 & 0.56 \\ [1pt]
&    density: n1$_{100}$ & -0.10 & 0.26 & -0.36 & 0.72 \\ 
&    dotplot: n1$_{100}$ & -0.77 & 0.26 & -3.00 & 0.00 & *** \\ 
&    histogram: n1$_{100}$ & -0.69 & 0.26 & -2.64 & 0.01 & ** \\ [1pt]
&    density:ratio & -0.28 & 0.47 & -0.59 & 0.56 \\ 
&    dotplot:ratio & -1.01 & 0.46 & -2.19 & 0.03 & * \\ 
&histogram:ratio & -0.28 & 0.47 & -0.59 & 0.55 \\ 
   \hline
\\[-5pt]
   \multicolumn{5}{l}{Signif. codes:  0 `***' 0.001 `**' 0.01 `*' 0.05 `.' 0.1}
\end{tabular}}
\end{center}
\caption{Results from a Generalized Linear Mixed Effects Model of Power of a lineup given design and all two-way interactions with effect size $d$, group size $n_1$, and ratio to second to first group size. Reference group for all effects are boxplot designs. The model was fit based on 2513 lineup evaluations of 208 participants. Note that estimate and error for all terms involving $n_1$ are multiplied by a factor of 100.}
\label{tbl:power2}
\end{table}

@@

Our third response variable is again the subjective assessment of confidence level. On average participants reported a confidence in their results of  3.41 (on a five point scale with 1=lowest and 5=highest), with a standard deviation of 0.149. Modeling confidence levels with a linear mixed effects model with the same structure as before resulted in only one significant effect beyond the intercept: participants report significantly higher confidence for increased number of  points in dotplot lineups. On average an increase of 100 points leads to an increase in confidence of 0.22 (with a standard error of 0.082). This finding is significant with a $p$-value of 0.007. 
Together with the previous results, we can tell that something interesting happens in the perception of dotplots as sample sizes increase. 
*** Follow-up experiments are planned. *** do we?

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=\linewidth]{conf-exp2} 
   \caption{example caption}
   \label{fig:example}
\end{figure}

>>>>

The two highly summarized designs benefit the most from an increase in sample size: both density charts and boxplots profit significantly more than histograms and dotplots. This might, again, be a hint towards a trade-off between visual complexity and amount of information shown as basis for decisions.

@@

Our third response variable is again the subjective assessment of confidence level. On average participants reported a confidence in their results of  3.41 (on a five point scale with 1=lowest and 5=highest), with a standard deviation of 0.149. Modeling confidence levels with a linear mixed effects model with the same structure as before resulted in only one significant effect beyond the intercept: participants report significantly higher confidence for increased number of  points in dotplot lineups. On average an increase of 100 points leads to an increase in confidence of 0.22 (with a standard error of 0.082). This finding is significant with a $p$-value of 0.007. 
Together with the previous results, we can tell that something worth of further investigation happens in the perception of dotplots as sample sizes increase.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=\linewidth]{conf-exp2} 
   \caption{Barcharts of marginal distributions of subjective confidence assessment by design. The distributions display prominent differences.}
   \label{fig:conf-margins}
\end{figure}

Figure \ref{fig:conf-margins} shows barcharts of confidence assessments by design. It is obvious that the distribution of confidence is very different from one design to the next. Boxplots and Dotplots show a similar pattern with an emphasis on high confidence levels, whereas histograms and density lineups have modes at a confidence level of three. One reason behind these differences in confidence distributions might be familiarity with the design - neither density charts nor histograms are particularly frequent displays for a general audience.
